3D Magnetic modelling of ellipsoidal bodies

### 2017-01-13 manuscript/manuscript.tex
methodology - geometrical aspects defining ellipsoids

\title{3D MAGNETIC MODELLING FOR ELLIPSOIDS}

@@

\affil[1]{Department of Geophysics, Observatorio Nacional, Rio de Janeiro, Brazil}
%\affil[]{}

@@

\correspondence{DIEGO TAKAHASHI TOMAZELLA (diego.takahashi@gmail.com)}

@@

\subsection{Geometrical aspects}

Let $(x, y, z)$ be a point referred to a Cartesian coordinate system with axes
$x$, $y$ and $z$ pointing to, respectively, North, East and down.

Consider an ellipsoidal body with centre at the point $(x_{c}, y_{c}, z_{c})$,
semi-axes defined by positive constants $a$, $b$, $c$, and orientation defined by
three angles $\alpha$, $\beta$, and $\gamma$.

The points $(x, y, z)$ located on the surface of this ellipsoidal body
satisfy the following equation:
\begin{equation}
(\mathbf{r} - \mathbf{r}_c)^T \mathbf{A} (\mathbf{r} - \mathbf{r}_c) = 1 \: ,
\label{eq:ellipsoid_surface}
\end{equation}
where $\mathbf{r} = [\begin{array}{ccc} x & y & z \end{array} ]^{\top}$,
$\mathbf{r}_{c} = [\begin{array}{ccc} x_{c} & y_{c} & z_{c} \end{array} ]^{\top}$,
$\mathbf{A}$ is a positive definite matrix given by
\begin{equation}
\mathbf{A} = \mathbf{V}
\left[ \begin{array}{ccc}
a^{-2} & 0 & 0 \\
0 & b^{-2} & 0 \\
0 & 0 & c^{-2} 
\end{array} \right] \mathbf{V}^{\top} \: ,
\label{eq:A}
\end{equation}
and $\mathbf{V}$ is an orthogonal matrix whose columns are defined
by unit vectors $\mathbf{v}_{1}$, $\mathbf{v}_{2}$, and $\mathbf{v}_{3}$.

The vectors $\mathbf{v}_{1}$, $\mathbf{v}_{2}$, and $\mathbf{v}_{3}$ 
have the same direction as the semi-axes $a$, $b$, $c$ of the 
ellipsoid, depend on the orientation angles 
$\alpha$, $\beta$, $\gamma$ and are defined 
according to the ellipsoid type.

For triaxial ellipsoids (i.e., $a > b > c$), the vectors 
$\mathbf{v}_{1}$, $\mathbf{v}_{2}$, and $\mathbf{v}_{3}$ are given
by \citep{clark1986magnetic}:
\begin{equation}
\mathbf{v}_{1} = \left[\begin{array}{c} 
                  -\cos\alpha \; \cos\delta \\
                  -\sin\alpha \; \cos\delta \\
                  -\sin\delta
                  \end{array} \right] \: ,
\label{eq:v1_triaxial_prolate}
\end{equation}
\begin{equation}
\mathbf{v}_{2} = \left[\begin{array}{c} 
                  \cos\alpha \; \cos\gamma \; \sin\delta + \sin\alpha \; \sin\gamma \\                   
                  \sin\alpha \; \cos\gamma \; \sin\delta - \cos\alpha \; \sin\gamma \\ 
                  -\cos\gamma \; \cos\delta
                  \end{array} \right] \: ,
\label{eq:v2_triaxial_prolate}
\end{equation}                   
\begin{equation}                    
\mathbf{v}_{3} = \left[\begin{array}{c} 
                  \sin\alpha \; \cos\gamma - \cos\alpha \; \sin\gamma \; \sin\delta \\                    
                  -\cos\alpha \; \cos\gamma - \sin\alpha \; \sin\gamma \; \sin\delta \\
                  \sin\gamma \; \cos\delta
                  \end{array} \right] \: .
\label{eq:v3_triaxial_prolate}
\end{equation}

Similarly, the vectors $\mathbf{v}_{1}$, $\mathbf{v}_{2}$, and $\mathbf{v}_{3}$
defining the semi-axes of prolate ellipsoids (i.e., $a > b = c$) are 
calculated by using equations \ref{eq:v1_triaxial_prolate}, \ref{eq:v2_triaxial_prolate}, 
and \ref{eq:v3_triaxial_prolate}, but with $\gamma = 0^{\circ}$ \citep{emerson1985magnetic}.

Finally, the vectors $\mathbf{v}_{1}$, $\mathbf{v}_{2}$, and $\mathbf{v}_{3}$
defining the semi-axes of oblate ellipsoids (i.e., $a < b = c$) are

PAREI AQUI

%whose eigenvectors define the 
%axes of the ellipsoid and the eigenvalues of A are the reciprocals of the squares of the semi-axes

>>>>

\title{3D Magnetic modelling of ellipsoidal bodies}

@@

\affil[1]{Department of Geophysics, Observat\'{o}rio Nacional, Rio de Janeiro, Brazil}
%\affil[]{}

@@

\correspondence{Diego Takahashi Tomazella (diego.takahashi@gmail.com)}

@@

\subsection{Geometrical parameters defining ellipsoids}

Let $(x, y, z)$ be a point referred to a Cartesian coordinate system with axes
$x$, $y$ and $z$ pointing to, respectively, North, East and down.
For convenience, we denominate this coordinate system as 
\textit{main coordinate system}.
Let us consider an ellipsoidal body with centre at the point $(x_{c}, y_{c}, z_{c})$,
semi-axes defined by positive constants $a$, $b$, $c$, 
where $a > b > c$, and orientation defined by
three angles $\alpha$, $\beta$, and $\gamma$.
The points $(x, y, z)$ located on the surface of this ellipsoidal body
satisfy the following equation:
\begin{equation}
(\mathbf{r} - \mathbf{r}_c)^T \mathbf{A} (\mathbf{r} - \mathbf{r}_c) = 1 \: ,
\label{eq:ellipsoid_surface}
\end{equation}
where $\mathbf{r} = [\begin{array}{ccc} x & y & z \end{array} ]^{\top}$,
$\mathbf{r}_{c} = [\begin{array}{ccc} x_{c} & y_{c} & z_{c} \end{array} ]^{\top}$,
$\mathbf{A}$ is a positive definite matrix given by
\begin{equation}
\mathbf{A} = \mathbf{V}
\left[ \begin{array}{ccc}
a^{-2} & 0 & 0 \\
0 & b^{-2} & 0 \\
0 & 0 & c^{-2} 
\end{array} \right] \mathbf{V}^{\top} \: ,
\label{eq:A}
\end{equation}
and $\mathbf{V}$ is an orthogonal matrix whose columns are defined
by unit vectors $\mathbf{v}_{1}$, $\mathbf{v}_{2}$, and $\mathbf{v}_{3}$.

The vectors $\mathbf{v}_{1}$, $\mathbf{v}_{2}$, and $\mathbf{v}_{3}$ 
depend on the orientation angles 
$\alpha$, $\beta$, $\gamma$ and are defined 
as follows \citep{clark1986magnetic}:
\begin{equation}
\mathbf{v}_{1} = \left[\begin{array}{c} 
                  -\cos\alpha \; \cos\delta \\
                  -\sin\alpha \; \cos\delta \\
                  -\sin\delta
                  \end{array} \right] \: ,
\label{eq:v1_triaxial_prolate}
\end{equation}
\begin{equation}
\mathbf{v}_{2} = \left[\begin{array}{c} 
                  \cos\alpha \; \cos\gamma \; \sin\delta + \sin\alpha \; \sin\gamma \\                   
                  \sin\alpha \; \cos\gamma \; \sin\delta - \cos\alpha \; \sin\gamma \\ 
                  -\cos\gamma \; \cos\delta
                  \end{array} \right] \: ,
\label{eq:v2_triaxial_prolate}
\end{equation}                   
\begin{equation}                    
\mathbf{v}_{3} = \left[\begin{array}{c} 
                  \sin\alpha \; \cos\gamma - \cos\alpha \; \sin\gamma \; \sin\delta \\                    
                  -\cos\alpha \; \cos\gamma - \sin\alpha \; \sin\gamma \; \sin\delta \\
                  \sin\gamma \; \cos\delta
                  \end{array} \right] \: .
\label{eq:v3_triaxial_prolate}
\end{equation}
For triaxial ellipsoids (i.e., $a > b > c$), the orthogonal matrix
$\mathbf{V}$ (equation \ref{eq:A}) is calculated by using equations
\ref{eq:v1_triaxial_prolate}, \ref{eq:v2_triaxial_prolate}, and
\ref{eq:v3_triaxial_prolate} as follows:
\begin{equation}
\mathbf{V} = \left[ \begin{array}{ccc}
\mathbf{v}_{1} & \mathbf{v}_{2} & \mathbf{v}_{3}
\end{array} \right] \: .
\label{eq:V_triaxial_prolate}
\end{equation}
Similarly, the matrix $\mathbf{V}$ (equation \ref{eq:A}) for
prolate ellipsoids (i.e., $a > b = c$) is calculated according
to equation \ref{eq:V_triaxial_prolate} by using equations
\ref{eq:v1_triaxial_prolate}, \ref{eq:v2_triaxial_prolate}, and
\ref{eq:v3_triaxial_prolate}, but with $\gamma = 0^{\circ}$ 
\citep{emerson1985magnetic}.
Finally, the matrix $\mathbf{V}$ (equation \ref{eq:A}) for
oblate ellipsoids (i.e., $a < b = c$) is calculated by 
using equations \ref{eq:v1_triaxial_prolate}, \ref{eq:v2_triaxial_prolate}, and
\ref{eq:v3_triaxial_prolate}, with $\gamma = 0^{\circ}$, as follows
\citep{emerson1985magnetic}:
\begin{equation}
\mathbf{V} = \left[ \begin{array}{ccc}
\mathbf{v}_{2} & \mathbf{v}_{1} & -\mathbf{v}_{3}
\end{array} \right] \: .
\label{eq:V_oblate}
\end{equation}
The orientation of the semi-axes $a$, $b$, and $c$ are defined 
by the first, second, and third columns of the matrix 
$\mathbf{V}$ given by equation \ref{eq:V_triaxial_prolate},
in the case of a triaxial or prolate ellipsoid, or the
matrix $\mathbf{V}$ given by equation \ref{eq:V_oblate},
in the case of an oblate ellipsoid.

The magnetic modelling of an ellipsoidal body is commonly performed
in a particular Cartesian coordinate system that is aligned 
with the body semi-axes
and has the origin coincident with the body centre.
For convenience, we denominate this particular coordinate 
system as \textit{local coordinate system}.
The relationship between the Cartesian coordinates 
$(\tilde{x}, \tilde{y}, \tilde{z})$ of a point in 
a local coordinate system and the Cartesian 
coordinates $(x, y, z)$ of the same point in the main
system is given by:
\begin{equation}
\tilde{\mathbf{r}} = \mathbf{V}^{\top} \left( \mathbf{r} - \mathbf{r}_{c} \right) \: ,
\label{eq:coord_transformation}
\end{equation}
where 
$\tilde{\mathbf{r}} = [\begin{array}{ccc} \tilde{x} & 
                                          \tilde{y} & 
                                          \tilde{z} \end{array} ]^{\top}$,
$\mathbf{r}$ and $\mathbf{r}_{c}$
are defined in equation \ref{eq:ellipsoid_surface} and the matrix $\mathbf{V}$
is defined according to equations \ref{eq:V_triaxial_prolate} or
\ref{eq:V_oblate}, depending on the ellipsoid type.

A last but not least set of parameters defining the geometry of
an ellipsoid are their focal points.
The focal points are defined on the axes of the local coordinate 
system as follows:
\begin{equation}
\left(\tilde{x}, \tilde{y}, \tilde{z} \right) = 
\begin{cases}
\left( \pm \sqrt{a^{2} - b^{2}}, 0, 0 \right) \\
\left( \pm \sqrt{a^{2} - c^{2}}, 0, 0 \right) \\
\left( 0, \pm \sqrt{b^{2} - c^{2}}, 0 \right)
\end{cases} \: .
\label{eq:focal-points}
\end{equation}
All ellipsoids sharing a common set of focal points form a family of confocal ellipsoids.

\subsection{Auxiliary variable $\lambda$}

The methods commonly applied in the magnetic modelling of
ellipsoidal bodies make use of an auxiliary variable $\lambda$,
which is a real number.

\subsection{Magnetic induction}
